### 2022-06-19 manuscript/manuscript.Rmd
major features added for gam models reporting and functionalities

---
title: ""
bibliography: "misc/ref.bib"
csl: "misc/vancouver.csl"
output: 
  word_document:
    reference_docx: "ref.docx"
---

**Título**: [título]

@@

**Autores**: Primer autor,^1^ Segundo autor,^2^ Tercer autor,^3^ Cuarto autor,^4^ Quinto autor,^5^ Sexto autor.^6^ \*

@@

## Resumen

**Objetivo**: [...]. **Material y métodos**: [...]. **Resultados**: [...]. **Conclusión**: [...].

**Palabras clave**: [...].

@@

# Introducción

@@

# Material y métodos

## Participantes

@@

## Instrumentos

### Instrumento 1

@@

### Instrumento 2

@@

## Procedimiento

@@

## Análisis estadístico

[...], using the *R* programming language for statistical computing, [@rlanguage].

@@

# Discusión

@@

# Conclusión

@@

# Agradecimientos

>>>>

---
title: ""
bibliography: "misc/ref.bib"
#csl: "misc/vancouver.csl"
output: 
  word_document:
    reference_docx: "misc/ref.docx"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  include = FALSE
)
```

@@

**Authors**: Primer autor,^1^ Segundo autor,^2^ Tercer autor,^3^ Cuarto autor,^4^ Quinto autor,^5^ Sexto autor.^6^ \*

@@

@@

# Introduction

@@

# Material y methods

## Participants

@@

## Instruments

### Instrument 1

@@

### Instrument 2

@@

## Procedures

@@

## Statistical analysis

Data is presented as median (*Mdn*) and interquartile range (*IQR*) for continuous variables, and as absolute and relative sample size for categorical/discrete variables.

A non-parametric approach was used since the underlying distribution of continuous outcomes, evaluated through analytical and graphical methods, did not follow a Gaussian distribution.

In order assess the differences in developmental scores between males and females, the *Wilcoxon* rank-sum test was used, meanwhile the chi-square test ($\chi^2$) was used to evaluate the goodness-of-fit ($\chi^2_{gof}$) and the independence of factors ($\chi^2_{Pearson}$).

Generalized additive models (GAM) were used to describe linear and non-linear relationships between developmental characteristics through smooth terms that are represented using penalized regression splines [@wood2011fast]. The restricted maximum likelihood estimation method was used for the selection of the smoothing parameter, and the thin-plate regression spline as the smoothing basis method, as they are the optimal smoother of any given basis dimension/rank [@wood2003thin]. GAMs were fitted using the mcgv package [@wood2017generalized]. To describe the smooth term, we use approximative derivative to summarise de trend in terms of linear segments.

[...], using the *R* programming language for statistical computing, [@rlanguage] and complementary R packages for model fitting, visualization and complementary computations [@wood2017generalized; @hadley2016ggplot2; @daniel2021see; @dominique2020estimation].

@@

```{r load-packages}
source(file = "manuscript/misc/setup.R")
```

```{r}
# Sample
n_sample <- dataset[, .N]
#> 234

# Sample by sex
n_sex <- dataset[j = .(stat = paste0(.N, " (", round(.N*100/n_sample, 1), "%)")), 
                 keyby = .(sex = sexo_paciente)]
#>    sex        stat
#> 1:   F  94 (40.2%)
#> 2:   M 140 (59.8%)

# Prop test on sex
prop_sex <- dataset[, table(sexo_paciente)] |> 
  chisq.test() |> 
  report_chi(effsize = TRUE)
#> [1] "$\\chi^2_{gof}$ (1) = 9.04, *p* = 0.003, $C_{Pearson}$ = 0.19, CI~0.95%~[0.09, 1]"

```

From a total of `r n_sample` subjects with congenital hypotonia, `r n_sex["F", stat]` were females and `r n_sex["M", stat]` males (`r prop_sex`). The developmental characteristics of the sample can be seen in [Table 1](#tab1).

```{r}
mod <- gam(comunicacion_total ~ s(edad_cronologica_meses), 
    data = dataset,
    method = "REML")

# Overall slope
gam_coef <- report_slope(mod, trend = c("edad_cronologica_meses"))

# Linear segments
modelbased::estimate_smooth(
  data = modelbased::estimate_relation(mod, length = 100),
  x = "edad_cronologica_meses"
)

plot_gam(dataset, edad_cronologica_meses, comunicacion_total, round_to = 5,
         "Chronological age (months)", "Communication score", method = "REML")
```

When modelling the effect of chronological age on developmental skills, we found an averaged negative marginal effect on the communication score (`r gam_coef`), whereas from 1 to 5.7 moths old

# Discusion

@@

# Conclusion

@@

# Acknowledgment
